Case for Support

### 2021-11-16 application/Je-S_sections.tex
11/16 changes Maneesh

Understanding the brain is a major scientific challenge of our era. To succeed, scientists must be able to explain how the rich repertoire of animal behaviour relates to the activity in neural circuits across different brain regions. This necessitates careful design and control of behavioural experiments wherein experimental animals (e.g. non-human primates, rodents, fish, insects) perform particular tasks while experimenters either record or manipulate neural activity in candidate neural circuits. 
Experiments in behavioural and brain science laboratories require software that integrates and controls hardware from multiple recording devices (video, electrodes for neural activity measurement, sensors), and analysis tools that can interpret large and complex behavioural and neural datasets.

@@

To understand the brain, scientists aim to explain how the rich repertoire of animal behaviour relates to the activity in neural circuits across different brain regions. This necessitates careful design and precise control of behavioural experiments wherein experimental animals perform particular tasks while experimenters either record or manipulate neural activity in candidate neural circuits. Such experiments require software that integrates and controls hardware from multiple recording devices (cameras, electrodes, sensors), and analysis tools that can interpret large and complex behavioural and neural datasets. The lack of standardised tools for experimental control and analysis across the brain-and-behaviour community is incurring major inefficiencies, with disproportionate time allocated to software engineering and writing analysis code, and efforts often replicated across laboratories. Similarly, disparate experimental and analysis protocols makes is almost impossible to replicate experiments between research groups.
Bonsai is an open-source reactive visual programming environment that naturally orchestrates complex networks of sensors and hardware into workflows. Bonsai has been adopted by hundreds of scientists worldwide to provide interactive experimental control in behavioural and brain sciences. Here we aim to extend Bonsai’s functionality with a toolbox of online and offline Machine Intelligence tools for analysis of behavioural and neural data (video-based analysis of behavioural motifs, real-time and off-line analysis of neural signals), and create an open-access platform for software sharing and integration with multiple programming languages. Bonsai’s enhanced functionality will enable new types of research, accelerate discovery and improve efficiency by (i) providing access to such tools for scientists lacking expertise, (ii) reducing duplication of effort, and (iii) standardising the data processing streams, thus increasing reproducibility across laboratories.
[12 chars over]

@@

>>>>

Understanding the brain is a major scientific challenge of our era. To succeed, scientists must be able to explain how animal behaviour relates to neural activity across different brain regions. This requires careful design and manipulation of behavioural experiments, where experimenters either record or manipulate neural activity while the animals (e.g. non-human primates, rodents, fish, insects) perform particular tasks.

Experiments in behavioural and brain science laboratories require software that integrates and controls hardware from multiple recording devices (video, electrodes for neural activity measurement, sensors), and analysis tools that can interpret large and complex behavioural and neural datasets.

@@

To understand the brain, scientists aim to explain how animal behaviour relates to the activity in neural circuits. This necessitates design and precise control of behavioural experiments wherein experimental animals perform particular tasks while experimenters either record or manipulate neural activity in candidate neural circuits. Such experiments require data acquisition software that integrates and controls hardware from multiple recording devices (cameras, electrodes, sensors), and analysis tools that can interpret large and complex behavioural and neural datasets. Progress is held back by the lack of standardised tools for rapid design and implementation of experimental protocols, and the difficulty of integrating state-of-the-art data processing and neuroinformatics into custom experimental designs. The fields of behavioural and brain sciences have consequently suffered from both inefficiency and poor reproducibility, due to disparate data acquisition and analysis solutions created independently across laboratories.   
To address these challenges, we propose  to  extend,  enhance,  maintain  and  support \textbf{Bonsai},  a  fully  integrated  software  environment  to enable cutting-edge reproducible laboratory systems neuroscience experiments in animal models, with a particular emphasis on machine-intelligence-enabled, real-time neuroinformatics methods. While Bonsai is already adopted by hundreds of scientists worldwide to provide interactive experimental control, we aim to extend Bonsai’s functionality with a toolbox of online and offline Machine Intelligence tools for analysis of behavioural and neural data (video-based analysis of behavioural motifs, real-time and off-line analysis of neural signals), and create an open-access platform for software sharing and integration with multiple programming languages. These to the Bonsai ecosystem will enable new types of research, increase user base, and dramatically improve efficiency and reproducibility of research across behavioural and brain sciences.
[44 chars over]

@@

Resource is Bonsai.
Monitor animal welfare

### 2021-11-16 application/backgroundToTheResource.tex
11/16 changes Maneesh

At present, progress in the field is held back by the challenge of rapidly designing and implementing new experimental protocols, and the difficulty of integrating state-of-the-art data processing and
neuroinformatics into custom experimental designs.
%
\textbf{Reproducibility}, in particular, depends on the availability of standardized tools for experimental control and analysis, shared between laboratories (Baker, 2016; Ioannidis, 2005). Although such tools exist in genetics, astronomy, physics and medicine (Fish et al., 2016; Abdalla et al., 2018, CERN Education, Communications and Outreach Group, 2018, Dickinson et al., 2016, Bycroft et al., 2018), they are mostly lacking in neuroscience, with documented consequences for the reliability of ``discoveries'' (Baker, 2016; Botvinik-Nezer et al., 2020; Button et al., 2013).

\setlength{\columnsep}{1em}
\begin{wrapfigure}{r}{0.5\textwidth}
  \includegraphics[width=0.5\columnwidth,viewport=65 50 500 800,clip]{figures/bonsai.pdf}
  \label{fig:bonsai}
\end{wrapfigure}

@@

@@

%

%

% This resource will be built on the platform established by \textbf{Bonsai}.
%

% an established
% experimental control software ecosystem with a long development history and a
% large user base in systems neuroscience (Figure~\ref{fig:bonsai}).

% \subsection{Bonsai}
% \label{sec:bonsai}
%
% We will 
% 
% 
% 
% A central goal of the software resource proposed here is to address this
% problem, by extending Bonsai~(an experimental control software ecosystem;
% Figure~\ref{fig:bonsai} and Section~\ref{sec:bonsai}) with state-of-the-art
% online (and offline) data analysis methods
% (Figure~\ref{fig:proposedBonsaiExtensions} and
% Section~\ref{sec:functionality}).
% Bonsai is unique amongst experimental control systems in that
% it allows users with no programming expertise to design and implement
% sophisticated experiments.
%
% The latest Its large and rapidly growing community of users demonstrates the demand for such software resources.
%

>>>>

At present, progress in the field is held back by the challenge of rapidly designing and implementing new experimental protocols, and the difficulty of integrating state-of-the-art data processing and
neuroinformatics into custom experimental designs.
%
\textbf{Reproducibility}, in particular, depends on the availability of standardised tools for experimental control and analysis, shared between laboratories (Baker, 2016; Ioannidis, 2005). Although such tools exist in genetics, astronomy, physics and medicine (Fish et al., 2016; Abdalla et al., 2018, CERN Education, Communications and Outreach Group, 2018, Dickinson et al., 2016, Bycroft et al., 2018), they are mostly lacking in neuroscience, with documented consequences for the reliability of ``discoveries'' (Baker, 2016; Botvinik-Nezer et al., 2020; Button et al., 2013).

\setlength{\columnsep}{1em}
\begin{wrapfigure}{r}{0.5\textwidth}
  \includegraphics[width=0.5\columnwidth,viewport=65 50 500 800,clip]{figures/bonsai.pdf}
  \label{fig:bonsai}
\end{wrapfigure}
\stepcounter{figure}

@@

Bonsai achieves robust behavioural control, hardware interaction and signal acquisition while ensuring remarkable ease of use.  The proposed project will provide valuable enhancements.  We will develop, test and deploy a new software workflow making it possible to incorporate external packages into Bonsai's C\#-based environment.  Building on this new extensibility, we will then incorporate a suite of machine-learning-driven analysis tools into the Bonsai environment.  These added capabilities will allow  video-based behavioural analysis as well as real-time interfacing of population-scale activity with laboratory control.  In parallel we will develop the infrastructure needed to evaluate competing data analytic approaches within a single experimental framework

@@

% A central goal of the software resource proposed here is to address this
% problem, by extending Bonsai~(an experimental control software ecosystem;
% Figure~\ref{fig:bonsai} and Section~\ref{sec:bonsai}) with state-of-the-art
% online (and offline) data analysis methods
% (Figure~\ref{fig:proposedBonsaiExtensions} and
% Section~\ref{sec:functionality}).
% Bonsai is unique amongst experimental control systems in that
% it allows users with no programming expertise to design and implement
% sophisticated experiments.
%
% The latest Its large and rapidly growing community of users demonstrates the demand for such software resources.
%

### 2021-11-16 application/narrativeCVSahani.tex
11/16 changes Maneesh

% 3) How have you contributed to research teams and the development of
% others?

% From QQR:
As a PI, I have supervised a total of \textbf{X} students and \textbf{X} postdocs.  
In the last five years, 4 of students have moved to postdoctoral positions (at
Aalto, Cambridge, CMU and Stanford), 1 works at DeepMind, and 1 holds a faculty offer at Shanghai Tech.  In all, 8 previous students have now secured faculty (or group leader) positions at Cambridge, CMU, Columbia, Copenhagen, Janelia (2), Pittsburgh and Shanghai.
%
Of the postdoctoral fellows I have supervised in the last five years, 1 is doing further postdoc work in Bonn; 1 holds an independent
Research Fellowship at Radboud; and 2 have faculty positions: at Edinburgh and at UBC.

@@

Gatsby directorship

% Examples might include: project management, supervision, mentoring or line management 
% contributions critical to the success of a team or team members, or where you exerted 
% strategic leadership in shaping the direction of a team, organisation, company or institution.

@@

I have been an active member of the academic community. I co-founded the Neural Coding, Computation and Dynamics meeting in 2007, and was in the organising committee in 2015 and 2017.

@@

I have sat on various grant award panels for the BMBF and DFG, and the (US) NSF/NIH collaborative research
in computational neuroscience (CRCNS) programme. I have also acted as a reviewer for proposals for the Bernstein prize, Einstein Foundation, EPSRC, MRC, Novo Nordisk Foundation (Denmark) and NWO (Netherlands).

>>>>

% 3) How have you contributed to research teams and the development of
% others? % Examples might include: project management, supervision, mentoring or line management 
% contributions critical to the success of a team or team members, or where you exerted 
% strategic leadership in shaping the direction of a team, organisation, company or institution.

% From QQR:
As a PI, I have supervised a total of \textbf{X} students and \textbf{X} postdocs.  
In the last five years, 4 of students have moved to postdoctoral positions (at
Aalto, Cambridge, CMU and Stanford), 1 works at DeepMind, and 1 holds a faculty offer at Shanghai Tech.  In all, 8 previous students have now secured faculty (or group leader) positions at Cambridge, CMU, Columbia, Copenhagen, Janelia (2), Pittsburgh and Shanghai.
%
Of the postdoctoral fellows I have supervised in the last five years, one is doing further postdoc work in Bonn; another holds an independent
Research Fellowship at Radboud; and 2 have faculty positions: at Edinburgh and at UBC.

@@

Since 2017, I have been Director of the Gatsby Computational Neuroscience Unit, and have spearheaded its new strategic research direction. I created two new roles to support and develop the new strategy, and have recruited 2 new faculty members who have added new and highly relevant research strands to our research portfolio. As Director, I sit in the Executive Leadership Committee of the Faculty of Life Sciences at UCL.

@@

Current collaborations/networks?

I have been an active member of the academic community by contributing the development of programme content for many scientific meetings, critically assessing the intellectual content of research proposals and awards applications, and reviewing the research and strategic direction of various world-class research institutions.

I co-founded the Neural Coding, Computation and Dynamics meeting in 2007, and was in the organising committee in 2015 and 2017.

@@

I have sat on various grant award panels for the BMBF and DFG (Germany), and the (US) NSF/NIH collaborative research
in computational neuroscience (CRCNS) programme. I have also acted as a reviewer for proposals for the Bernstein prize, the Einstein Foundation, EPSRC, MRC, Novo Nordisk Foundation (Denmark) and NWO (Netherlands).
